---
title: "HW4"
author: "Duc Le"
date: "9/24/2020"
output:
  html_document:
    df_print: paged
---

### Problem 1
```{r}
library(DAAG)
library(dplyr)

data(possum)
attach(possum)
```
```{r}
hist1 = hist(age)
hist2 = hist(age, breaks = seq(0,9,1.5))

```
The difference we see in hist2 with the custom bins is due to the fact that we attempt to split a discrete variable, age, into float bins. 
```{r}
plot(density(age, na.rm = T))
```
Histograms can be useful and easier to interpret discrete data and its frequency. KDE's on the other hand could come in handy when we need to visualize continuous data. 

### Problem 2
```{r}
hist(earconch)
boxplot(earconch~sex)
```
The length of ear conch does seem to differ by sex. The median length of females is much higher than males'. Also, the males' data in the 75 percentile seems to be more spread out vs. the females' 75 percentile.

### Problem 3
```{r}
data = select(possum, -case, -site, -Pop, -sex)
plot(data)
plot(skullw, hdlngth)
points(x=mean(skullw),y=mean(hdlngth),col="red",pch=17,cex=1.5)
```

