---
title: "HW3"
author: "Duc Le"
date: "9/20/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
Loading the data frame created from task 5.
```{r}
library(tibble)
library(dplyr)

load("map.coords.RData")
ccA = read.csv("ccA.csv")
ccB = read.csv("ccB.csv")
gdp = read.csv("gdp.csv")
df = merge(ccA, ccB, all = TRUE)
df2 = cbind(df, gdp)
```

### Problem 1
Inner merge
```{r}
inner = merge(df2, coords, all = FALSE)
dim(inner)
```
Left merge
```{r}
left = merge(df2, coords, all.x = TRUE)
dim(left)
```
Right merge
```{r}
right = merge(df2, coords, all.y = TRUE)
dim(right)
```

There's a clear difference in the # of rows after each merge. This is obviously caused by the number of unmatched rows merged between the data frames, depending on each method.

### Problem 2

2a)
```{r}
data = inner %>%
  select(gdp, lat)

quantiles = quantile(data$gdp)
quantiles
```

2b)
```{r}
gdp.q = findInterval(data$gdp, quantiles, rightmost.closed = TRUE)
table(gdp.q)
```
Yes the function returns 4 quartiles & their associated medians. It also provides the amount of elements in each quartile.

2c)
```{r}
rev_gdp.q = factor(gdp.q,levels = 4:1)
table(rev_gdp.q)
```

2d)
```{r}
q1 = data %>%
  filter(gdp <= quantiles[2])
q1.lat = mean(abs(q1$lat))
q1.lat

q2 = data %>%
  filter(gdp >= quantiles[2] & gdp <= quantiles[3])
q2.lat = mean(abs(q2$lat))
q2.lat

q3 = data %>%
  filter(gdp >= quantiles[3] & gdp <= quantiles[4])
q3.lat = mean(abs(q3$lat))
q3.lat

q4 = data %>%
  filter(gdp >= quantiles[4] & gdp <= quantiles[5])
q4.lat = mean(abs(q4$lat))
q4.lat
```
There isn't a big discrepancy between the means of latitude associated to each quartile. 
