---
title: "HW3"
author: "Duc Le"
date: "12/7/2020"
output: pdf_document
---

### Problem 1
```{r}
library(dplyr)
library(MASS)
library(VGAM)
library(mlbench)

rm(list = ls())
d = get(data(BostonHousing))
str(d)
q = quantile(d$medv, c(0.25, 0.75))

d$medv = ifelse(d$medv < q[1], 0, ifelse(d$medv >= q[1] & d$medv <= q[2], 1, 2))
table(d$medv)

model1 = vglm(medv~., multinomial(refLevel = 1), data = d)
summary(model1)
```
Looking at the summary of the coefficients, we can eliminate variables that aren't statistically
significant to simplify the model such as: crim, zn, indus, cha1. Since my reference level = 1,
these variables don't contribute much to the log odds of being between the 25th & 75th percentile
Similarly, they aren't significant when it comes to predicting the log odds of being above the
75th percentile.

In addition, we can view the effects of these coefficients the same way we did in normal logistic regression, with the exception that each Beta's will have different effects on the odds depending on what you selected for the reference level. 

### Problem 2
```{r}
model2 = vglm(medv~., family = propodds(), data = d)
summary(model2)
lrtest(model1, model2)
```
Using the likelihood ratio statistical test, we are able to reject H0 due to the statistical significance, thus the model from part a is correct.