---
title: "HW1 Biostats"
output:
  html_document: default
  pdf_document: default
---

### Problem 1 

**1a)**
H0: m1 = m2 & Ha: m1 != m2.
Null: both teaching methods produce the same results.
Alt: the two teaching methods produce different results.
```{r}
m1 = c(79,66,57,91,42,59)
m2 = c(71,43,58,78,20,56)
alpha = 0.01

normality = function(data){
  shapiro = shapiro.test(data)
  p.value = shapiro$p.value
  if(p.value <= alpha){print("Reject the Null. Normality is not met")}
  else if(p.value > alpha){print("Do not reject the Null. Normality is met")}
} 

normality(m1)

normality(m2)
```

**1b)** Under the assumption that all 12 students are different, we can use the t-test.
The function below will conduct a t-test with 99% confidence level.
```{r}
t_test_99ci = function(x, y){
  t_test_result = t.test(x, y, conf.int=TRUE, conf.level=0.99)
  p.value = t_test_result$p.value
  if(p.value <= alpha){print("Reject the Null")}
  else if(p.value > alpha){print("Do not reject the Null")}
}

t_test_99ci(m1, m2)
```
The conducted t-test gives us a 99% confidence level to not reject the null hypothesis given the resulted p-value.


**1b)** H0: m1 = m2. Ha: m1 != m2.

Since the students were tested twice, we can use the paired t-test. The function below will conduct a paired t-test with 99% confidence level.

```{r}
paired_t_test = function(x,y,alpha){
  t_test_result = t.test(x, y, conf.int=TRUE, conf.level=0.99, paired = T)
  p.value = t_test_result$p.value
  if(p.value <= alpha){print("Reject the Null")}
  else if(p.value > alpha){print("Do not reject the Null")}
}

paired_t_test(m1, m2, alpha)
```
Null hypothesis states that method 1 yields the same improvement as method 2. If the same students are tested twice, the paired t-test shows that the p value is not sufficient enough to reject the null hypothesis.


**1c)**
```{r}
conf.test = t.test(m1, mu = 80, conf.level = 0.9)
conf.test$conf.int
```
The 90% confidence interval is [51.42, 79.91]. Ha: true mean != 80.

### Problem 2

```{r}
library(foreign)
library(dplyr)
wcgs = read.dta("wcgs.dta")
g0 = wcgs %>%
  filter(typchd69 == 0)
g1 = wcgs %>%
  filter(typchd69 == 1)
g2 = wcgs %>% 
  filter(typchd69 == 2)
g3 = wcgs %>%
  filter(typchd69 == 3)
```

**2a)**
```{r}
alpha = 0.05

normality(g0$weight)
normality(g1$weight)
normality(g2$weight)
normality(g3$weight)
```
Normality is not met for any of the group.


**2b)**
```{r}
bartlett_test = function(v1.v2, data){
  result = bartlett.test(v1.v2, data)
  p = result$p.value
  
  if(p < alpha){print("Reject the Null. Variances are different across groups")}
  else if(p >= alpha){print("Cannot reject the Null. Same variances across groups")}
}

bartlett_test(weight~typchd69, wcgs)
```


**2c)**
```{r}
bonferroni = function(x, g){
  test = pairwise.t.test(x, g, p.adjust.method = "bonf")
  p = test$p.value
  print(p)
}

anovaTest = function(v1.v2, data){
  test = anova(lm(v1.v2, data))
  p = test$'Pr(>F)'[1]
  if(p < alpha){print("Reject the Null. There's no difference of means")}
  else if(p >= alpha){print("Cannot reject the Null.")}
}

anovaTest(weight~typchd69, wcgs)
bonferroni(wcgs$weight, wcgs$typchd69)
```


**2d)**
```{r}
kruskal = function(v1.v2, data){
  test = kruskal.test(v1.v2, data)
  p = test$p.value
  if(p < alpha){print("Reject the Null. There's no difference of medians")}
  else if(p >= alpha){print("Cannot reject the Null.")}
}

kruskal(weight~typchd69, wcgs)
```
The Kruskal-Wallis test rejects the null. We apply the Bonf. adj. Wilcoxon signed rank test.


```{r}
wilcox.rank = function(v1, v2, data){
  test = wilcox.test(v1, v2, p.adj = "bonf")
  p = test$p.value
  if(p < alpha){print("Reject the Null. The 4 groups have the same distributions.")}
  else if(p >= alpha){print("Cannot reject the Null.")}
}

wilcox.rank(wcgs$weight, wcgs$typchd69, wcgs)
```

### Problem 3

**3a)** H0: m0 = m1. Null Hypothesis: diet isn't effective. Ha: m0 != m1. Alt Hypothesis: diet is effective.

```{r}
diet = read.delim("diets.txt",sep = "" , header = T)
diet1 = diet %>%
  filter(Diet == 1)
diet2 = diet %>%
  filter(Diet == 2)
diet3 = diet %>%
  filter(Diet == 3)
diet4 = diet %>% 
  filter(Diet == 4)
```

**3b)**

Preliminary tests for ANOVA:

Normality test using Shapiro-Wilk
```{r}
normality(diet1$WeightLoss)
normality(diet2$WeightLoss)
normality(diet3$WeightLoss)
normality(diet4$WeightLoss)
```
Normality is met for all diet groups.

Testing for difference of variances, means, medians & distributions:
```{r}
bartlett_test(WeightLoss~Diet, diet)

anovaTest(WeightLoss~Diet, diet)

kruskal(WeightLoss~Diet, diet)

wilcox.rank(diet$WeightLoss, diet$Diet, diet)
```


**3d)**
```{r}
bonferroni(diet$WeightLoss, diet$Diet)
```
We reject the null since all p-values < alpha.


**3f)**
```{r}
diet.aov = aov(WeightLoss~Diet, diet)
summary(diet.aov)
```
The p.value after using aov() is ~0.0003 < alpha. Thus we reject H0.


**3g)**
```{r}
t = pairwise.t.test(diet$WeightLoss, diet$Diet)
t$p.value
```
We reject the null since all p-values < alpha.

### Problem 4
```{r}
run = "aabbaaaabbbaababbaaabbbbbababbbabbb"

letters = strsplit(run,"")[[1]]
count = table(letters)
A = count[1][[1]]
B = count[2][[1]]

i = 1
runs = 1
len = length(letters)

while (i < (len - 1)){
  if (letters[i] != letters[i+1]) {runs = runs + 1}
  i = i + 1
}
runs
```
The runs is even so we can use the even probability function.

```{r}
prob = (2*(choose(A-1,runs/2 -1))*choose(B-1, runs/2 - 1))/choose(A+B, A)
prob
if (prob > alpha){print("We can assume randomness")}
```

